The Social Conditioning of Fear

### 2023-12-06 scripts/manuscript.qmd
Intro

Historically, despite all the aforementioned contexts and individual-level characteristics being identified as salient to individuals’ self-reported fear at night, they were usually studied in isolation. Notably, scholars have tended to focus either on individual-level characteristics, such as race and sexual orientation, or on physically defined contexts like schools and neighborhoods and social contexts like networks, families, with little consideration of the joint contributions of multiple contexts and multilevel effects simultaneously. Instead, we hypothesize that, similar to "the social determinants of health" [@krieger2001], there exist the social determinants of fear--intersectional social positionalities impact the likelihood of individuals's sense of threat from social surroundings, i.e., fear at night. This is an exploratory project for a dissertation chapter idea that one of the co-authors wants to pursue. With descriptive statistics, we are interested in investigating the inequalities in terms of fear of walking alone at night among US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education.

>>>>

Historically, despite all the aforementioned contexts and individual-level characteristics being identified as salient to individuals’ self-reported fear at night, they were usually studied in isolation. Notably, scholars have tended to focus either on individual-level characteristics, such as race and sexual orientation, or on physically defined contexts like schools and neighborhoods and social contexts like networks, families, with little consideration of the joint contributions of multiple contexts and multilevel effects simultaneously. Instead, we hypothesize that, similar to "the social determinants of health" [@krieger2001], there exist the social determinants of fear--intersectional social positionalities impact the likelihood of individuals' sense of threat from social surroundings, i.e., fear at night.

This is an exploratory project for a dissertation chapter idea that one of the co-authors wants to pursue. With descriptive statistics, we are interested in investigating the inequalities in terms of fear of walking alone at night among US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education. We use the 2022 data from the General Social Survey (GSS) and the data science tools from EDLD 651 class to delineate the data distribution and basic correlations between "fear of walkign alone at night" and our other variables of interest.
